Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-09-16 writedown/period.noneq.msm/paper.tex
backup

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibirium MD -- this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^T(t)\cdot \vect p(t)
\end{align}
where $\vect L^T(t)$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$, $p_i(t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p_i(0)\ge 0$ componentwise, then $p_i(t)\ge 0$ and $\sum_i
p_i(t) = \sum_ip_i(0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^T(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^T(t)$ is a
stochastic matrix satisfying $\myphi^T(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^T\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Temporal discretization: Floquet theory}
\label{sec:floquet}

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. It satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L^T(t+T) \vect Y(t)=\vect L^T(t)\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T)$.  
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
and for the probability at integral period we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0),
\end{align}
where $\myphi^m(T)$ denotes the $m$th power of the matrix $\myphi(T)$.
Since $\myphi(T)$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially coomplex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with left eigenvector $\vect e=(1,\ldots,1)^T$ and a right eigenvector $\mymu$
satisfying
$\myphi(T)\mymu=\mymu$.
From now on, we assume $\myphi(T)$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). In this case $\mymu$ is the stationary measure in the sense that
\begin{align}
\label{eq:floq-14}  
\myphi(mT) \mymu = \mymu,\quad m=0,1,2,\ldots,
\end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
\begin{align}
\label{eq:floq-15}  
\myphi(mT)\vect p(0)\to \mymu,\quad m\to\infty,
\end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.

Using the Floquet theorem, the time-inhomogeneous Markov jump process $X_t$
is simplified into a \emph{time-homogeneous} (not necessarily
reversible) Markov jump process $\tilde X_{m} = X_{mT}, \ m\in\mathbb
N$.
Therefore, we prefer to consider the discrete-time process $\tilde X_{m}$ instead of the time-continuous process $X_t$
because the powerful theories and
computational tools for time-homogeneous Markov processes can be directly applied. Many of these tools require the transition matrix $\myphi(T)$ to satisfy the detailed balance condition. The computations in the Appendix show that $\myphi$ will in general \emph{not} satisfy this condition; in fact  the deviation form reversibility can be estimated from the work of the periodic driving does to the system.

There is no doubt that information within one period is lost by using this temporal
discretization, however, information regarding the long-term behavior of the system on time-scales
much longer than the period will be perfectly described because of $\tilde X_{m} = X_{mT}\approx x_{mT}$ whenever our spatial discretization is fine enough.
At the same time,
the computational cost of generating $\tilde X_{m}$ is much less demanding
than the brute force simulations of NEMD, which implies lower
statistically uncertainty in calculating the observables of interest.

@@

% \subsection{Building the Markov State Model from the Floquet transition matrix}
If the discretization cells $\Omega_i$, $i=1,\ldots,N$ form a fine
partition of the molecular state space, the Markov chain defined via
the transition matrix 
\begin{equation}\label{P}
\vect P=\myphi^T(T)
\end{equation}
 is discrete in time but in space it still is a fine-scale description of the transport properties of the
dynamics with a very large number $N$ of states.  Now we want to coarsen our description much further by
constructing a Markov State Model (MSM) for $\vect P$ with $K\ll N$
\emph{macrostates} that should be the metastable states of the system: The resulting $K\times K$ MSM transition matrix $\hat{\vect P}$
then defines the coarse grained long term kinetics that shall
approximate the original long term kinetics well. 
The idea behind MSM building is that given the molecular system under consideration exhibits metastable conformations then it is usually possible to construct a
relatively small number of discrete sets --the metastable sets that form the so-called macrostates-- that
correctly describe the slow  dynamics, and 
in each set the fast dynamics relaxes on some timescale significantly shorter than the metastable timescales.
Then if the MSM dynamics reproduces the slow timescales and the corresponding transitions
of the original dynamics~\eqref{eq:disc-1},
the former is considered to be a good approximation of the later.

@@

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing} which allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $K$ core sets $C_1,\ldots, C_K\subset S$ that are appropriate metastable sets. 
Following~\cite{A19-31,schuette2011markov} the process $(\tilde X_m)$ associated with $P$ is coarse grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ just has $K$ states associated with the sets $C_j$, $j=1,\ldots,K$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(\tilde X_m)$, i.e., trajectories of $(\tilde X_m)$ induce trajectories of $(\hat{X}_m)$: We set $\hat{X}_m=j$ if the last core set that the process $(\tilde X_m)$ entered prior to or at time $m$ has been the core set $C_j$.
\end{itemize}
Now consider infinitely long trajectory of $(\tilde X_m)$. Because of ergodicity we know that the states in this trajectory will be distributed due to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory we can consider the probability $q_j^\bwd(i)$ that conditioned on $\tilde X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(\tilde X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\[
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j|X_m=i),
\]
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
\recheck{COMMENT: Here it is a \emph{conditional} probability,
  due to the time-homogeneity,
  it does not matter whether it is defined from quasi-stationary probability
  or not.}
From the last equation we get that the stationary distribution of the milestone process is given by
\[
\hat{\mu}_j=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\]
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}_j$.

@@

To illustrate how the discretization works in practice we take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied in Ref.~\cite{wang2014exploring}.
The system was simulated in a $2.7\times 2.7\times 2.7$~$\textrm{nm}^3$ periodic simulation
region, with one alanine dipeptide molecules described by the CHARMM27 force field~\cite{foloppe2000all} dissolved in 641 TIP3P water molecules~\cite{jorgensen1983comparison}.
The grid-based energy correction map (CMAP)~\cite{mackerell2004extending} was used
to correct the backbone dihedral angle energies.
All simulations were performed by a home-modified
Gromacs~4.6.5~\cite{pronk2013gromacs} with CHARMM27 force field implemented~\cite{bjelkmar2010implementation}.
The alanine dipeptide was put into the local thermostating
environment, with a spherical dynamical region of radius~1.0~nm
centered at the alpha-carbon.
The Langevin thermostat with target temperature $\mathcal T = 300$~K
and time-scale $\tau_T = 0.1$~ps was
coupled to the thermostated region.
The whole system was coupled to the Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} (in standard Gromacs implementation) with $\tau_P = 2.0$~ps to
keep the system at 1~Bar. The non-equilibirium trajectories
were integrated by the Leap-frog scheme with a time-step of 0.002~ps.
The short-range van der Waals interactions were cut-off at 1.00~nm, and were smoothed from
0.95~nm to 1.00~nm by the ``\texttt{shift}'' method provide by Gromacs.
The energy conserving 
Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm} method (``\texttt{pme-switch}'') was
used to compute the long-range electrostatic interaction,
with the same real-space cut-off radius as the van der Waals interactoins,
and the Gromacs default Fourier spacing of 0.12~nm.
The splitting
parameter optimized with respect to the electrostatic force computing
accuracy by Gromacs tool \texttt{g\char`_pme\char`_error}~\cite{wang2010optimizing}.
The neighbor list was updated every 5 time-steps with a list-building radius 1.20~nm.
All hydrogen involving covalent bonds were constrained by the LINCS algorithm~\cite{hess1997lincs}, except the water molecules that were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
The whole system was driven by a periodic electric field
$E(t) = E_0\sin(2\pi t/T)$ and $D(x) = (1,0,0)^T$
with intensity of the field being $E_0 = 1.0$~V/nm and period being
$T = 10$~ps.
The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
% \redc{Write a lot of details on NEMD.}
The equilibrium configurations were prepared by an equilibrium MD simulation
of $10^6$~ps long, along which the snapshots were saved every 50~ps.
% In the
% equilibrium simulation the system was coupled to a global Langevin thermostat
% with coupling time-scale $\tau_T = 0.5$~ps.
The branching
trajectories were each 4000~ps long, and the system reached
non-equilibrium quasi-stationary state in roughly 300~ps.

@@

Based on a given spatial discretization we can aggregate the transition matrix $\vect P=\myphi^T(T)$ just by counting the transition behavior of MD trajectories 
\[
\vect P_{ij}=\prob\left(\vect X_T=j\mid \vect X_0=i \right).
\]
However, $\vect P$ allows to approximate the original dynamics on multiples $mT$ of the period only. In order to have a time-continuos description we need the generator $L(t)$ of the Master equation. 
If the discretization is fine enough one possible approximation to $L(t)$ is via
the following forward finite difference scheme:
\begin{align}
  \label{eqn:tmp4}
  L_{ij}(t) \approx \frac{1}{\tau}
  \,[\, \prob (\vect X_{t+\tau} = j \mid \vect X_{t} = i) - \delta_{ji} \,],
  \quad i,j\in S
\end{align}
and $\tau$ is an appropriate small enough lag-time.
In the following 
we investigate the discretization
quality with respect to  the choice of $N_\dih$ and  lag-time $\tau$.

@@

We estimate the discretized generator $L(t), \ t\in[0,T)$
from NEMD trajectories generated by $\ml(t)$ in different time intervals $[t_1,
t_2]$.
Whenever the discretization is fine enough the discretized dynamics approximates the original
dynamics well, and the time-periodic generator $L(t)$ should not depend on the choice of the interval $[t_1,t_2]$ in the estimation procedure (\ref{eqn:tmp4}), provided that the initial state of the system
is not very far from the stationary state at long-time limit.
Therefore, this is an indicator for calibrating the discretization quality.
We compute $L(t)$ by two discretizations $N_\dih=2$ and $N_\dih=20$, and two
choices of time intervals $[0, 80]$~ps and $[320, 400]$~ps, and then
compare the time-dependent probability $\prob\big(\phi_t\in[0,180), \psi_t\in [0,180)\big)$
with the (brute force) NEMD result in Fig.~\ref{fig:tmp2} using a lag-time $\tau=0.5$~ps.
Using $N_\dih=2$ the dynamics depends on the time interval used for
calculating the generator: using time interval $[0, 80]$~ps the discretized
dynamics deviates from the NEMD result,
while using time interval $[320, 400]$~ps the discretized dynamics can only
reproduces the NEMD result after 300~ps.  This therefore indicates poor 
approximations to the original dynamics with $N_\dih=2$. The reason is that the
discretization with $N_\dih=2$ is too coarse so that the dynamics cannot be fully
equilibriated within the lag-time $\tau$ in each discretized set,
therefore, the discretization presents state dependency.  For
$N_\dih=20$, the discretized dynamics does not depend on the time interval of
calculating the generator, and is consistent with the
NEMD simulation within the error bar. Therefore, throughout this paper we use $N_\dih=20$
to discretize  the dihedral angle space of alanine dipeptide.
%For a good statistical accuracy, if not stated otherwise,
%we will use the full trajectories, i.e.~a time interval of
%$[0,4000]$~ps for estimating the discretized generator $L(t)$.

@@

\section{Reversibility of the original dynamics}
\label{sec:revs}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibirium MD -- this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^{\top}(t)\cdot \vect p(t)
\end{align}
where $\vect L^{\top}(t)$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$, $p_i(t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p_i(0)\ge 0$ componentwise, then $p_i(t)\ge 0$ and $\sum_i
p_i(t) = \sum_ip_i(0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^{\top}(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^{\top}(t)$ is a
stochastic matrix satisfying $\myphi^{\top}(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^{\top}\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Temporal discretization: Floquet theorem}
\label{sec:floquet}

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. It satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L^{\top}(t+T) \vect Y(t)=\vect L^{\top}(t)\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T)$.
\recheck{This is known as the Floquet theorem~\cite{floquet1883equations}.}
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
and for the probability at integral period we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0),
\end{align}
where $\myphi^m(T)$ denotes the $m$th power of the matrix $\myphi(T)$.

\recheck{
Using the Floquet theorem, the time-inhomogeneous Markov jump process $X_t$
is therefore discretized into a \emph{time-homogeneous} (not necessarily
reversible) Markov jump process $\tilde X_{m} = X_{mT}, \ m\in\mathbb
N$, which is generated by transition matrix
\begin{equation}\label{P}
\vect P=\myphi^{\top}(T).
\end{equation}
We prefer to consider the discrete-time process $\tilde X_{m}$ instead of the time-continuous process $X_t$
because the powerful theories and
computational tools for time-homogeneous Markov processes can be directly applied. It worth noting that many of these tools require the transition matrix $P$ to satisfy the detailed balance condition. The computations in the Appendix~\ref{sec:app-revs} show that $P$ will in general \emph{not} satisfy this condition; in fact  the deviation form reversibility can be estimated from the work of the periodic driving does to the system.
There is no doubt that information within one period is lost by using this temporal
discretization, however, information regarding the long-term behavior of the system on time-scales
much longer than the period will be perfectly described because of $\tilde X_{m} = X_{mT}\approx x_{mT}$ whenever our spatial discretization is fine enough.
At the same time,
the computational cost of generating $\tilde X_{m}$ is much less demanding
than the brute force simulations of NEMD, which implies lower
statistically uncertainty in calculating the observables of interest.
}

Since \recheck{$P$} \recheck{(\underline{COMMENT}: using $P$ instead of $\myphi(T)$.)} is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially coomplex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with \recheck{right} eigenvector $\vect e=(1,\ldots,1)^{\top}$ and a \recheck{left} eigenvector $\mymu$
satisfying
\recheck{$\mymu^{\top}P=\mymu^{\top}$.}
From now on, we assume $P$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). \recheck{In this case $\mymu$ is the stationary measure in the sense that
$
 \mymu^{\top} P^m = \mymu^{\top},\ m\in\mathbb N
$,
% \begin{align}
% \label{eq:floq-14}  
% \end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
$
\vect p^{\top}(0)\,P^m \to \mymu^{\top},\ m\to\infty,
$
% \begin{align}
% \label{eq:floq-15}  
% \end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.}

@@

% \subsection{Building the Markov State Model from the Floquet transition matrix}
If the discretization cells $\Omega_i$, $i=1,\ldots,N$ form a fine
partition of the molecular state space, the Markov chain defined via
the transition matrix $P$
 is discrete in time, but in space it still is a fine-scale description of the transport properties of the
dynamics with a very large number $N$ of states.  Now we want to coarse our description much further by
constructing a Markov State Model (MSM) for $\vect P$ with $K\ll N$
\emph{macrostates} that should be the metastable states of the system: The resulting $K\times K$ MSM transition matrix $\hat{\vect P}$
then defines the coarse grained long term kinetics that shall
approximate the original long term kinetics well. 
The idea behind MSM building is that given the molecular system under consideration exhibits metastable conformations then it is usually possible to construct a
relatively small number of discrete sets --the metastable sets that form the so-called macrostates-- that
correctly describe the slow  dynamics, and 
in each set the fast dynamics relaxes on some timescales significantly shorter than the metastable timescales.
Then if the MSM dynamics reproduces the slow timescales and the corresponding transitions
of the original dynamics~\eqref{eq:disc-1},
the former is considered to be a good approximation of the later.

@@

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing} which allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $K$ core sets $C_1,\ldots, C_K\subset S$ that are appropriate metastable sets. 
Following~\cite{A19-31,schuette2011markov} the process $(\tilde X_m)$ associated with $P$ is coarse grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ just has $K$ states associated with the sets $C_j$, $j=1,\ldots,K$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(\tilde X_m)$, i.e., trajectories of $(\tilde X_m)$ induce trajectories of $(\hat{X}_m)$: We set $\hat{X}_m=j$ if the last core set that the process $(\tilde X_m)$ entered prior to or at time $m$ has been the core set $C_j$.
\end{itemize}
Now consider infinitely long trajectory of $(\tilde X_m)$. Because of ergodicity we know that the states in this trajectory will be distributed due to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory we can consider the probability $q_j^\bwd(i)$ that conditioned on $\tilde X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(\tilde X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\begin{align}
\label{eq:msm-bwd-committor}  
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j \mid X_m=i),
\end{align}
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
\recheck{\underline{QUESTION}: Here it is a \emph{conditional} probability,
  due to the time-homogeneity,
  it does not matter whether it is defined from quasi-stationary probability
  or not.}
From the last equation we get that the stationary distribution of the milestone process is given by
\begin{align}  
\hat{\mu}_j=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\end{align}
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}_j$.

@@

To illustrate how the discretization works in practice we take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied in Ref.~\cite{wang2014exploring}.
The system was simulated in a $2.7\times 2.7\times 2.7$~$\textrm{nm}^3$ periodic simulation
region, with one alanine dipeptide molecules described by the CHARMM27 force field~\cite{foloppe2000all} dissolved in 641 TIP3P water molecules~\cite{jorgensen1983comparison}.
The grid-based energy correction map (CMAP)~\cite{mackerell2004extending} was used
to correct the backbone dihedral angle energies.
All simulations were performed by a home-modified
Gromacs~4.6.5~\cite{pronk2013gromacs} with CHARMM27 force field implemented~\cite{bjelkmar2010implementation}.
The alanine dipeptide was put into the local thermostating
environment, with a spherical dynamical region of radius~1.0~nm
centered at the alpha-carbon.
The Langevin thermostat with target temperature $\mathcal T = 300$~K
and time-scale $\tau_T = 0.1$~ps was
coupled to the thermostated region.
The whole system was coupled to the Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} (in standard Gromacs implementation) with $\tau_P = 2.0$~ps to
keep the system at 1~Bar. The non-equilibirium trajectories
were integrated by the Leap-frog scheme with a time-step of 0.002~ps.
The short-range van der Waals interactions were cut-off at 1.00~nm, and were smoothed from
0.95~nm to 1.00~nm by the ``\texttt{shift}'' method provide by Gromacs.
The energy conserving 
Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm} method (``\texttt{pme-switch}'') was
used to compute the long-range electrostatic interaction,
with the same real-space cut-off radius as the van der Waals interactoins,
and the Gromacs default Fourier spacing of 0.12~nm.
The splitting
parameter optimized with respect to the electrostatic force computing
accuracy by Gromacs tool \texttt{g\char`_pme\char`_error}~\cite{wang2010optimizing}.
The neighbor list was updated every 5 time-steps with a list-building radius 1.20~nm.
All hydrogen involving covalent bonds were constrained by the LINCS algorithm~\cite{hess1997lincs}, except the water molecules that were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
The whole system was driven by a periodic electric field
$E(t) = E_0\sin(2\pi t/T)$ and $D(x) = (1,0,0)^{\top}$
with intensity of the field being $E_0 = 1.0$~V/nm and period being
$T = 10$~ps.
The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
% \redc{Write a lot of details on NEMD.}
The equilibrium configurations were prepared by an equilibrium MD simulation
of $10^6$~ps long, along which the snapshots were saved every 50~ps.
% In the
% equilibrium simulation the system was coupled to a global Langevin thermostat
% with coupling time-scale $\tau_T = 0.5$~ps.
The branching
trajectories were each 4000~ps long, and the system reached
non-equilibrium quasi-stationary state in roughly 300~ps.

@@

Based on a given spatial discretization we can aggregate the transition matrix $\vect P=\myphi^{\top}(T)$ just by counting the transition behavior of MD trajectories 
\[
\vect P_{ij}=\prob\left(\vect X_T=j\mid \vect X_0=i \right).
\]
However, $\vect P$ allows to approximate the original dynamics on multiples $mT$ of the period only. In order to have a time-continuos description we need the generator $L(t)$ of the Master equation. 
If the discretization is fine enough one possible approximation to $L(t)$ is via
the following forward finite difference scheme:
\begin{align}
  \label{eqn:tmp4}
  L_{ij}(t) \approx \frac{1}{\tau}
  \,[\, \prob (\vect X_{t+\tau} = j \mid \vect X_{t} = i) - \delta_{ji} \,],
  \quad i,j\in S
\end{align}
and $\tau$ is an appropriate small enough lag-time \recheck{(small error in
  the finite difference approximation).}
\recheck{
  % COMMENT: shouldn't be discussions on the lag-time and the Markovianity?
  In the meanwhile,
  since the dimensionality is reduced by using only a few collective
  variables, the lag-time should be chosen large enough so that
  the original dynamics $x_t$ is properly relaxed at
  the unresolved degrees of freedoms in timescales shorter than the lag-time
  (possible if the collective variables capture the slow dynamics).
  Then the implied Markovianity assumption in the master equation is
  reasonable, and the discretization from $\ml(t)$ to $L(t)$ is state-independent.}
In the following 
we investigate the discretization
quality with respect to  the choice of $N_\dih$ and  lag-time $\tau$.

\recheck{\underline{QUESTION}: Is the above discussion correct?
  Does the ``Markovianity'' always need a time discretization with the
  lag-time?
}

@@

We estimate the discretized generator $L(t), \ t\in[0,T)$
from NEMD trajectories generated by $\ml(t)$ in different time intervals $[t_1,
t_2]$.
\recheck{As discussed above, }whenever  the discretized dynamics approximates the original
dynamics well, and the time-periodic generator $L(t)$ should not depend on the choice of the interval $[t_1,t_2]$ in the estimation procedure (\ref{eqn:tmp4}), provided that the initial state of the system
is not very far from the stationary state at long-time limit.
Therefore, this is an indicator for calibrating the discretization quality.
We compute $L(t)$ by two discretizations $N_\dih=2$ and $N_\dih=20$, and two
choices of time intervals $[0, 80]$~ps and $[320, 400]$~ps, and then
compare the time-dependent probability $\prob\big(\phi_t\in[0,180), \psi_t\in [0,180)\big)$
with the (brute force) NEMD result in Fig.~\ref{fig:tmp2} using a lag-time $\tau=0.5$~ps.
Using $N_\dih=2$ the dynamics depends on the time interval used for
calculating the generator: using time interval $[0, 80]$~ps the discretized
dynamics deviates from the NEMD result,
while using time interval $[320, 400]$~ps the discretized dynamics can only
reproduces the NEMD result after 300~ps.  This therefore indicates poor 
approximations to the original dynamics with $N_\dih=2$. The reason is that the
discretization with $N_\dih=2$ is too coarse so that the dynamics cannot be fully
equilibriated within the lag-time $\tau$ in each discretized set,
therefore, the discretization presents state dependency.  For
$N_\dih=20$, the discretized dynamics does not depend on the time interval of
calculating the generator, and is consistent with the
NEMD simulation within the error bar. Therefore, throughout this paper we use $N_\dih=20$
to discretize  the dihedral angle space of alanine dipeptide.
%For a good statistical accuracy, if not stated otherwise,
%we will use the full trajectories, i.e.~a time interval of
%$[0,4000]$~ps for estimating the discretized generator $L(t)$.

@@

\section{Reversibility of the original dynamics}
\label{sec:app-revs}
